Relict deciduous oaks show high resilience to drought despite the effect of land-use legacies

### 2019-02-06 man/figures/evi_vs_severe_droughts.Rmd
update figures

---
title: "Growing season vs. Drought"
author: "AJ Perez-Luque (@ajpelu)"
date: "2017 Feb"
output:  
  md_document:
    variant: gfm
bibliography: ../refs/references.bib
csl: ../refs/ecology.csl
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,warning=FALSE, message=FALSE)
```

```{r, warning=FALSE, message=FALSE}
library("tidyverse")
library("here")
library("lubridate")
library("anytime")
library("rdryad")
```

```{r}
# Datos de iv 
iv <- read.csv(here::here("data/evi/iv_composite.csv"), header = TRUE, sep = ',')

@@

```{r}
# datos de Drought
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)

@@

```{r}
micolor <- '#455883'

@@

```{r}
# PLOT EVI + DROUGHT 
combinado2 <- ggplot(evi_profile_dat, aes(cd, y=mean)) + 
  scale_x_date(labels = function(x) format(x, "%b"), breaks = date_breaks('month')) + 
  theme_bw() +
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank()) +
  theme_classic() +
  geom_segment(data = dy, aes(y = mean, x= iD, xend =eD, yend = mean, colour = d_severity), 
               size = 8, alpha = 0.7) +
  geom_text(data = dy, aes(label = year, x = (iD + floor(eD-iD)/2))) +
  scale_colour_gradient(low = "orange", high = "brown") + 
  labs(colour = "Drought Severity") +
  geom_errorbar(aes(ymin = mean - 10*se, ymax= mean + 10*se), width=4, colour=micolor, size=.8) + 
  geom_line(colour=micolor, size=.8) + 
  geom_point(size=3,colour=micolor) +
  geom_point(size=1.5, colour='white') + 
  ylab('EVI') + xlab('Date')

pdf(here::here('/out/drought_evi', 'evi_droughts.pdf'), width=8, height = 5)
combinado2
dev.off()

# Get data from Dryad 
# Peres de Lis 2016 et al. https://datadryad.org/resource/doi:10.5061/dryad.1cn19
# x <- dryad_files('10.5061/dryad.1cn19')

>>>>

---
output:
  word_document
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
library("lubridate")
library("anytime")
library("rdryad")
library("scales")
```

**Figure S7.** EVI annual profile (average of the period 2000-2016) for *Q. pyrenaica* forests in Sierra Nevada and drought events. Horizontal bars correspond to the most severe droughts for Sierra Nevada since 1900 (computed as in Table S3). Their position indicate the start and end months of each drought event. Bars length show the duration of the drought event (number of consecutive months with SPEI lower than -1.28, see [@Pascoa2017]).

```{r preparedata}
# Datos de iv 
iv <- read.csv(here::here("data/evi/iv_composite.csv"), header = TRUE, sep = ',')

@@

# datos de Drought
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)

@@

```{r plotEVI, eval = FALSE, echo = FALSE}
micolor <- '#455883'

@@

```{r}
# PLOT EVI + DROUGHT 
combinado2 <- ggplot(evi_profile_dat, aes(cd, y=mean)) + 
  scale_x_date(labels = function(x) format(x, "%b"), breaks = date_breaks('month')) + 
  theme_bw() +
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank()) +
  theme_classic() +
  geom_segment(data = dy, aes(y = mean, x= iD, xend =eD, yend = mean, colour = d_severity), 
               size = 8, alpha = 0.7) +
  geom_text(data = dy, aes(label = year, x = (iD + floor(eD-iD)/2))) +
  scale_colour_gradient(low = "orange", high = "brown") + 
  labs(colour = "Drought Severity") +
  geom_errorbar(aes(ymin = mean - 10*se, ymax= mean + 10*se), width=4, colour=micolor, size=.8) + 
  geom_line(colour=micolor, size=.8) + 
  geom_point(size=3,colour=micolor) +
  geom_point(size=1.5, colour='white') + 
  ylab('EVI') + xlab('')

# Get data from Dryad 
# Peres de Lis 2016 et al. https://datadryad.org/resource/doi:10.5061/dryad.1cn19
# x <- dryad_files('10.5061/dryad.1cn19')
```

```{r plot_out, fig.height=5, dpi = 300, out.width = 0.8}
combinado2
```

### 2019-02-06 man/tables/t_droughts.Rmd
update figures

##### 
**Appendix S3.** Drought events for Sierra Nevada based on SPEI index (12 months scale). A drought event starts in the month when SPEI falls below the threshold of -1.28 [@Pascoa2017]. A drought event is considered only when SPEI value are below threshold for at least two consecutive months [*e.g.* @Spinoni2015; @Spinoni2017a; @Pascoa2017]. The ***duration*** of a drought event is the number of consecutive months with the SPEI lower than a certain threshold. ***Severity*** of a drought event is the sum of the SPEI values (absolute values) during the duration of the drought event. ***Intensity*** and ***Lowest SPEI*** refer to the mean and lowest value of SPEI respectively during the drought event duration. SPEI data were obtained for all 0.5º grid cells covering Sierra Nevada.

```{r}
# peores sequias 
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)
dy <- dy %>% arrange(desc(d_duration))
# solamente aquellos mayor de 2 meses 
dys <- dy %>% 
  filter(d_duration > 2) %>%  
  mutate(year = ifelse(minyear == maxyear, 
                       minyear, 
                       paste0(minyear, '-', maxyear))) %>% 
  dplyr::select(-minyear, -maxyear, -cat, -month_peak) %>% 
  as.data.frame()

>>>>

##### 
**Table S3.** Drought events for Sierra Nevada based on SPEI index (12 months scale). A drought event starts in the month when SPEI falls below the threshold of -1.28 [@Pascoa2017]. A drought event is considered only when SPEI values are below the threshold for at least two consecutive months [*e.g.* @Spinoni2015; @Spinoni2017a; @Pascoa2017]. The ***duration*** of a drought event is the number of consecutive months with the SPEI lower than a certain threshold. ***Severity*** of a drought event is the sum of the SPEI values (absolute values) during the duration of the drought event. ***Intensity*** and ***Lowest SPEI*** refer to the mean and lowest value of SPEI respectively during the drought event duration. SPEI data were obtained for all 0.5º grid cells covering Sierra Nevada.

```{r}
# peores sequias 
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)
dy <- dy %>% arrange(desc(d_duration)) %>% 
  arrange(desc(d_severity))
# solamente aquellos mayor de 2 meses 
dys <- dy %>% 
  filter(d_duration > 2) %>%  
  mutate(year = ifelse(minyear == maxyear, 
                       minyear, 
                       paste0(minyear, '-', maxyear))) %>% 
  dplyr::select(-minyear, -maxyear, -cat, -month_peak) %>% 
  as.data.frame()

### 2019-02-06 man/tables/t_robust_anovas.Rmd
update figures

**Table 3.** Robust two-way ANOVAs of the resilience metrics of greenness (EVI) and tree-growth (BAI) for the two factors drought events (in 2005 and 2012) and site.

@@

dff_bai <- bai %>% gather(temp, numeros, F, p) %>% 
  unite(temp1, variables, temp) %>% 
  spread(temp1, numeros) %>% 
  mutate(factores = recode(factores, 
                           "disturb_year" = "Drought event", 
                           "disturb_year:site" = "Drought event:Site", 
                           "site" = "Site")) %>% 
  arrange(match(factores, c("Drought event", "Site", "Drought event:Site"))) %>% 
  mutate_at(c(2,4,6,8), round, 1) %>% 
  mutate_at(c(3,5,7,9), round, 3) %>% 
  mutate(variable = 'BAI')

>>>>

**Table 3.** Robust two-way ANOVAs of the resilience metrics of greenness (EVI) and tree-growth (BAI) for the two drought events (in 2005 and 2012) and site.

@@

dff_bai <- bai %>% gather(temp, numeros, F, p) %>% 
  unite(temp1, variables, temp) %>% 
  spread(temp1, numeros) %>% 
  mutate(factores = recode(factores, 
                           "disturb_year" = "Drought event", 
                           "disturb_year:site" = "Drought event*Site", 
                           "site" = "Site")) %>% 
  arrange(match(factores, c("Drought event", "Site", "Drought event*Site"))) %>% 
  mutate_at(c(2,4,6,8), round, 1) %>% 
  mutate_at(c(3,5,7,9), round, 3) %>% 
  mutate(variable = 'BAI')

### 2019-02-06 man/tables/t_sampled_plots.Rmd
update figures

**Table 1.** Characteristics of sampled plot. Lat = latitude; Long = longitude. Dbh and height of all trees, Basal Area (BA), Density and SRD (Size ratio proportional to distance) are computed for all trees within a 10-m radius of focal trees (see methods). Temp.: annual average of minimun and maximum temperatures. Values shown here correspond to site averages. Standard deviations are shown in parentheses. Different letters indicate statistically significant differences between sites (Kruskal-Wallis test followed by Dunn’s test, p<0.05).

@@

ta <- ta %>% set_header_labels(name = "Site", lat_m = "Lat (\u00B0)", long_m = "Long (\u00B0)", 
                               elev = "Elevation\n(m)", slope = "Slope (\u00B0)",  prec = "Prec.\n(mm)",
                               temp = "Temp.\n(\u00B0 C)", n_trees = "# trees\n(# cores)", dn = "Dbh (cm)",
                               h = "Height\n(m)", age = "Age\n(years)", dn_all = "Dbh all\n(cm)", 
                               h_all = "Height all\n(m)", ba = "BA\n(m2/ha)", 
                               stand_density = "Density\n(trees/ha)", srd = "SRD") %>% 
  add_header(name="", lat_m ="", long_m="", elev="", slope ="", prec = "", temp = "",
             n_trees = "Cored trees", dn = "Cored trees", h = "Cored trees", age = "Cored trees", 
             dn_all = "Competence", h_all = "Competence", ba = "Competence", 
             stand_density = "Competence", srd = "Competence", top = TRUE) %>% 
  merge_at(i = 1, j =c(8:11) , part = "header") %>% 
  merge_at(i = 1, j =c(12:16) , part = "header") %>% 
  align(align = "center", part = "header") %>% 
  bold(part = 'header') %>% 
  fontsize(i= 1, part = "header", size = 8) %>% 
  fontsize(i= 2, part = "header", size = 8) %>% 
  italic(i=1, part = "header") %>% 
    # General format 
  fontsize(part = "body", size = 8) %>% 
  autofit(add_w = 0) %>% 
  align(align = "center") %>% 
  width(j = c(2,3), width = 0.35) %>% 
  width(j = c(6,7,8), width = 0.6)

>>>>

**Table 1.** Characteristics of sampled plot. Lat = latitude; Long = longitude. Dbh and height of all trees, Basal Area (BA), Density and SRD (Size ratio proportional to distance) are computed for all trees within a 10-m radius of focal trees (see methods). Temp.: annual average of mean monthly minimun and maximum temperatures. Values shown here correspond to site averages. Standard deviations are shown in parentheses. Different letters indicate statistically significant differences between sites (Kruskal-Wallis test followed by Dunn’s test, p<0.05).

@@

ta <- ta %>% set_header_labels(name = "Site", lat_m = "Lat (\u00B0)", long_m = "Long (\u00B0)", 
                               elev = "Elevation\n(m)", slope = "Slope (\u00B0)",  prec = "Prec.\n(mm)",
                               temp = "Temp.\n(\u00B0 C)", n_trees = "# trees\n(# cores)", dn = "Dbh (cm)",
                               h = "Height\n(m)", age = "Age\n(years)", dn_all = "Dbh all\n(cm)", 
                               h_all = "Height all\n(m)", ba = "BA\n(m2/ha)", 
                               stand_density = "Density\n(trees/ha)", srd = "SRD") %>% 
  add_header(name="", lat_m ="", long_m="", elev="", slope ="", prec = "", temp = "",
             n_trees = "Cored trees", dn = "Cored trees", h = "Cored trees", age = "Cored trees", 
             dn_all = "Competition", h_all = "Competition", ba = "Competition", 
             stand_density = "Competition", srd = "Competition", top = TRUE) %>% 
  merge_at(i = 1, j =c(8:11) , part = "header") %>% 
  merge_at(i = 1, j =c(12:16) , part = "header") %>% 
  align(align = "center", part = "header") %>% 
  bold(part = 'header') %>% 
  fontsize(i= 1, part = "header", size = 8) %>% 
  fontsize(i= 2, part = "header", size = 8) %>% 
  italic(i=1, part = "header") %>% 
    # General format 
  fontsize(part = "body", size = 8) %>% 
  autofit(add_w = 0) %>% 
  align(align = "center") %>% 
  width(j = c(2,3), width = 0.35) %>% 
  width(j = c(6,7,8), width = 0.6)
